---
title: "Assignment 3 CO2 Emissions Valuation"
author: "Hannah Irish, Kristin Art, Lillian Liu, Nadine Snyder"
date: "2023-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(forecast)
```

Read in historical emissions data for industrial and transportation emissions
```{r}

##read in historical data on emissions from CARB
ghg_historical_ca <- read_csv(here::here("data/arb_ghg_2000-2020.csv")) %>%
  janitor::clean_names()

##separate transportation
trans_hist_ghg_ca <- ghg_historical_ca %>%
  filter(sector == "Transportation") 

##separate industrial
ind_hist_ghg_ca <- ghg_historical_ca %>%
  filter(sector == "Industrial")
```


Use linear regression to find the trend of emissions, no intervention
```{r}

##linear model for transportation
trans_emissions_lm <- lm(mmt_co23 ~ year, data = trans_hist_ghg_ca) 

##extract slope
trans_coef <- coef(trans_emissions_lm)
trans_int <- trans_coef[1] 
trans_slope <- trans_coef[2]

##linear model for industrial
ind_emissions_lm <- lm(mmt_co23 ~ year, data = ind_hist_ghg_ca) 

##extract slope
ind_coef <- coef(ind_emissions_lm)
ind_int <- ind_coef[1] 
ind_slope <- ind_coef[2]

```

Develop counterfactual emissions
```{r}
##define lm functions
calculate_ind_co2e <- function(year){
  co2e = (ind_slope*year + ind_int)*1000 
}

calculate_trans_co2e <- function(year){
  co2e = (trans_slope*year + trans_int)*1000
}

##make vector of years
year <- seq(2021, 2050) 

##turn into df and add columns for transportation and industrial emissions
counterfactual <- year %>% 
  as.data.frame() %>% 
  rename("year" = 1) %>% 
  mutate(transportation = calculate_trans_co2e(year),
         industry = calculate_ind_co2e(year))

## pivot longer
counterfactual_longer <- counterfactual %>% 
  filter(year <= "2050") %>% 
  pivot_longer(cols = c(2,3), names_to = "sector", values_to = "mt_co2e")

##plot to check
ggplot() + geom_line(data = counterfactual_longer, aes(x = year, y = mt_co2e, col = sector)) 
```


Read in Base Case Buildout Information
```{r}

##read in Reed paper buildout by year (estimated from graph)
h2_buildout <- read_csv(here::here("data/H2_from_electrolysis.csv")) %>%
  select(1,4)

colnames(h2_buildout) = c("year", "no_add")

##lag by 4 years for additionality scenario
h2_scenarios <- h2_buildout %>%
  mutate(add = lag(no_add,2)) #is this a 2-year lag?

##add 0 to first two rows of additionality
h2_scenarios$add[1] = 0
h2_scenarios$add[2] = 0

h2_scenarios_tidy <- h2_scenarios %>%
  pivot_longer(cols=c(2,3), names_to = "scenario", values_to = "mt_h2")

```

Read in industry and transportation (non-LDV) percentage values

```{r}

##read in data with the percentage of hydrogen demand attributable to non-LDVs and industry 
percent_demand <- read_csv(here::here("data/non-LDV_vs_Industry_Demand.csv")) %>%
  select(c(1,8,9))

```

Final H2 scenarios attributable to each sector
```{r}

##put in percent data
percents_df <- merge(percent_demand, h2_scenarios) 


##calculate h2 production that will go to each demand type
final_h2 <- percents_df %>%
  mutate("no_additionality_trans" = transportation_pct*no_add, 
         "no_additionality_ind" = industry_pct*no_add,
         "additionality_trans" = transportation_pct*add,
         "additionality_industry" = industry_pct*add) %>%
  select(c(1,6,7,8,9))


## choose additionality situations
additionality <- final_h2 %>%
  select(1, 4,5 ) 

colnames(additionality) =  c("year","transportation", "industry")

##calculate Co2equivalent saved by the hydrogen production for each end use and scenario

##1 kg h2 = 0.9gal diesel. 1 metric ton  1000kg. 1 gal diesel = 10.18*10^-3 CO2 tons so
#h2mt * 1000kg/mt * 0.9gal diesel/1kg h2 *10*10^-3 co2 kg/1gal diesel * 1ton/1000kg co2
additionality_co2 <- additionality %>%
  mutate("transportation" = transportation * 1000 *0.9*10.180 *0.001/1000)%>%
  ## 1 kg h2 = 0.000117 mcf natural gas *0.0551 ton CO2/1mcf natural gas
  #h2mt * 1000kg/mt * 0.000117 mcf nat gas/kgh2 * 0.0551 co2 tons/mcf natural gas
  mutate("industry" = industry *1000*0.000117*0.0551)

# additionality_tidy <- additionality_co2 %>%
#   pivot_longer(cols = c(2,3), names_to = "scenario", values_to = "mmt")

## subtract the saved emissions from the counterfactual scenario
additionality_co2 <- additionality_co2 %>% 
  left_join(counterfactual, by = "year") %>% 
  mutate(transportation = transportation.y - transportation.x,
         industry = industry.y - industry.x) %>% #calculate remaining emissions from counterfactual - saved emissions for additionality scenario
  select(year, transportation, industry)


## choose non=additionality situations
no_additionality <- final_h2 %>%
  select(1, 2,3 ) 

colnames(no_additionality) =  c("year","transportation", "industry")

##calculate Co2equivalent saved by the hydrogen production for each end use and scenario
no_additionality_co2 <- no_additionality %>%
  mutate("transportation" =  transportation * 1000 *0.9*10.180 *0.001/1000)%>%
  ## 1 kg h2 = 0.000117 mcf natural gas *0.0551 ton CO2/1mcf natural gas
  #h2mt * 1000kg/mt * 0.000117 mcf nat gas/kgh2 * 0.0551 co2 tons/mcf natural gas
  mutate("industry" = industry *1000*0.000117*0.0551)

# no_additionality_tidy <- no_additionality_co2 %>%
#   pivot_longer(cols = c(2,3), names_to = "scenario", values_to = "mmt")

## subtract the saved emissions from the counterfactual scenario
no_additionality_co2 <- no_additionality_co2 %>% 
  left_join(counterfactual, by = "year") %>% 
  mutate(transportation = transportation.y - transportation.x,
         industry = industry.y - industry.x) %>% #calculate remaining emissions from counterfactual - saved emissions for additionality scenario
  select(year, transportation, industry)


```


get monetary value (SCC)
```{r}

##get monetary value by using social cost of carbon
scc = 185 ##Using scc of 185 from RFF (based on Costello's recommendation)

##discount value
rho = 1/1.05 ##using dsct value of 5%

##add column for t and then calculate pv of costs for each scenario
counterfactual_scc <- counterfactual %>%
  mutate("industry" = industry*185, "transportation" = transportation*185) %>%
  filter(year %in% c("2024", "2026", "2028", "2030", "2032", "2034", "2036", "2038", "2040", "2042", "2044", "2046", "2048", "2050")) %>% #2023 is timestep 0, 2024 is timestep 1, filter to years that match other dfs
  mutate(total = industry+transportation, ##sum total cost for industry and transportation
         t = seq(1, 28, 2), #2023 is timestep 0, 2024 is timestep 1
         pv = (scc*total)/ rho^t) 


##same for additionality and no additionality scenarios
no_add_scc <- no_additionality_co2 %>%
  mutate("industry" = industry*185, "transportation" = transportation*185)%>%
  filter(year>= "2024") %>%
  mutate(total = industry+transportation,
         t = seq(1, 28, 2), #since values are for every 2 years, make timesteps same
         pv = (scc*total)/ rho^t)

add_scc <- additionality_co2  %>%
  mutate("industry" = industry*185, "transportation" = transportation*185) %>%
  filter(year >= "2024") %>%
  mutate(total = industry+transportation,
         t = seq(1, 28, 2), #since values are for every 2 years, make timesteps same
         pv = (scc*total)/ rho^t)

##sum costs for net present value
# npv_counterfactual = sum(counterfactual_scc$pv)
# npv_additionality = npv_counterfactual - sum(add_scc$pv) 
# npv_no_additionality = npv_counterfactual - sum(no_add_scc$pv)
# 
# ##look at results
# print(npv_counterfactual)
# print(npv_additionality)
# print(npv_no_additionality)

##kristin's rambling thoughts
#I added a step that subtracts the saved emissions from h2 from the emissions in the counterfactual to end up with actual emissions under the scenarios. Then, calculated the np cost based on the scc. Results below show that the scenario with h2 and no additionality is the least cost option, although it's pretty dang close to scenario with additionality. Both interventions are ~half the cost of the counterfactual, so woo!

npc_counterfactual = sum(counterfactual_scc$pv)
npc_additionality = sum(add_scc$pv) 
npc_no_additionality = sum(no_add_scc$pv)

min(npc_counterfactual,npc_additionality, npc_no_additionality)

print(npc_counterfactual) #max cost
print(npc_additionality) 
print(npc_no_additionality) #this is the minimum cost by a slight amount

```

